---
title: "My R Markdown Primer"
author: "Michael Phommavong"
date: "2025-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document provides a brief introduction to R Markdown, a framework that lets you combine text, code, and mathematical notation in a single reproducible report. R Markdown automatically runs your R code and displays results within the final output.


## What R Markdown Is

R Markdown is a powerful tool within the R ecosystem that allows users to create 
dynamic, reproducible documents by combining narrative text, R code, and the results 
(like tables, plots, statistical outputs) into a single file.


## LaTex math expression

Here is an inline math expression:
The standard deviation is computed as $\sigma = \sqrt{\frac{1}{n}\sum(x_i - \bar{x})^2}$.

Here is a displayed (centered) LaTeX equation:
$$
f(x) = \frac{1}{\sqrt{2\pi\sigma^2}} e^{-\frac{(x-\mu)^2}{2\sigma^2}}
$$


## Example Code 1: Loading a dataset
```{r}
# Load the built-in 'mtcars' dataset
data(mtcars)

# Show the first few rows
head(mtcars)
```


## Example Code 2: Summary statistics for mpg and basic scatter plot
```{r}
# Summary statistics for mpg
summary(mtcars$mpg)

# Basic scatter plot
plot(mtcars$hp, mtcars$mpg,
main = "MPG vs Horsepower",
xlab = "Horsepower",
ylab = "Miles per Gallon",
pch = 19)
```


## Reflection

Working with R Markdown has given me more experience working with code and learning how narrative text can be integrated into one document. By using tools such as "knit" in R Markdown, I can compile all the information into a formatted report and make analyses more reproducible.
